---
title: "R Notebook"
output: html_notebook
---

The [R plugin](https://www.jetbrains.com/help/pycharm/r-plugin-support.html) for IntelliJ-based IDEs provides
handy capabilities to work with the [R Markdown](https://www.jetbrains.com/help/pycharm/r-markdown.html) files.
To [add](https://www.jetbrains.com/help/pycharm/r-markdown.html#add-code-chunk) a new R chunk,
position the caret at any line or the code chunk, then click "+".

The code chunk appears:
```{r}
library(readxl)
df1 <- read_excel('C:\\Users\\ADMIN\\RProjects\\Time_Series\\Gross Sale Revenues - HSG.xlsx')
df1
```

Type any R code in the chunk, for example:
```{r}
gross_sale_revenue <- ts(df1, start = c(2008,1), frequency = 4)
time <- seq(gross_sale_revenue)
summary(time)
```
```{r}
s1 <- rep(c(1,0,0,0), 14)
s1 <- c(s1, c(1,0,0))

s2 <- rep(c(0,1,0,0), 14)
s2 <- c(s2, c(0,1,0))

s3 <- rep(c(0,0,1,0), 14)
s3 <- c(s3, c(0,0,1))

s4 <- rep(c(0,0,0,1), 14)
s4 <- c(s4, c(0,0,0))
```
```{r}
time_predict <- seq(60, 118)
data_predict <- data.frame(time = time_predict, seas2 = s3 , seas3= s4, seas4= s1)
data_predict
```
```{r}
## linear trend + seasonality
## additive form
reg_trend_seas_add <- lm(gross_sale_revenue ~ time + s2 +s3 + s4)
summary(reg_trend_seas_add)
```
```{r}
predict(reg_trend_seas_add, data_predict)
```
```{r}
## MAPE:full
MAPE( fitted(reg_trend_seas_add), df1$Gross_Sale_Revenues)
## MAPE:2022
MAPE( fitted(reg_trend_seas_add)[57:59], df1$Gross_Sale_Revenues[57:59])
```
```{r}
## linear trend + seasonality
## Multiplicative form
reg_trend_seas_mul <- lm(gross_sale_revenue ~ time + time * s2 + time * s3 + time * s4)
summary(reg_trend_seas_mul)
```
```{r}
predict(reg_trend_seas_mul, data_predict)
```
```{r}
## MAPE:full
MAPE( fitted(reg_trend_seas_mul), df1$Gross_Sale_Revenues)
## MAPE:2022
MAPE( fitted(reg_trend_seas_mul)[57:59], df1$Gross_Sale_Revenues[57:59])
```
```{r}
## Holt_Winter
## Holt_Winter_add
hw_add <- HoltWinters(gross_sale_revenue, seasonal = 'a')
hw_add
```
```{r}
predict(hw_add, n.ahead = 5)
```
```{r}
## MAPE:full
MAPE( fitted(hw(gross_sale_revenue, seasonal = "additive")), df1$Gross_Sale_Revenues)
## MAPE:2022
MAPE( fitted(hw(gross_sale_revenue, seasonal = "additive"))[57: 59], df1$Gross_Sale_Revenues[57:59])
```
```{r}
hw_mul <- hw(gross_sale_revenue, seasonal = 'm')
hw_mul
```
```{r}
predict(hw_mul, n.ahead = 5)
```
```{r}
MAPE( fitted(hw(gross_sale_revenue, seasonal = "m")), df1$Gross_Sale_Revenues)
## MAPE:2022
MAPE( fitted(hw(gross_sale_revenue, seasonal = "m"))[57: 59], df1$Gross_Sale_Revenues[57:59])
```

Now, click the **Run** button on the chunk toolbar to [execute](https://www.jetbrains.com/help/pycharm/r-markdown.html#run-r-code) the chunk code. The result should be placed under the chunk.
Click the **Knit and Open Document** to build and preview an output.
