---
title: "R Notebook"
output: html_notebook
---

The [R plugin](https://www.jetbrains.com/help/pycharm/r-plugin-support.html) for IntelliJ-based IDEs provides
handy capabilities to work with the [R Markdown](https://www.jetbrains.com/help/pycharm/r-markdown.html) files.
To [add](https://www.jetbrains.com/help/pycharm/r-markdown.html#add-code-chunk) a new R chunk,
position the caret at any line or the code chunk, then click "+".

The code chunk appears:
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(Hmisc)
library(tidyverse)
library(Metrics)
library(astsa)
library(tseries)
library(readxl)
library(tsutils)
library(forecast)
library(urca)
library(lmtest)
library(vars)
library(tsDyn)
```

Type any R code in the chunk, for example:
```{r}
df <- read_excel("C:\\Users\\Admin\\RProjects\\Time_Series\\GSO_2005_2019_example.xlsx", col_names = TRUE)
df
```
```{r}
summary(df$v07)
```
```{r}
v07s <- (df$v07 - mean(df$v07))/ sd(df$v07)
median(v07s)
```
```{r}
cor(df$v07, df$v08)
```
```{r}
# test for correlation of v07 and v06, at 5%
cor.test(df$v07, df$v06)
```
p-value =  2.2e-16 < 5% -> Reject H0, correlated
```{r}
#Compare correlation among v03, v05, vo6, v07
cor(data.frame(df$v03, df$v05, df$v06, df$v07))
```
## Generate time trend variable 1
# Regress v07 on time series
coefficients of determination = ?
```{r}
t <- seq_along(df$v07)
plot.ts(df$v07)
```
```{r}
summary( lm(df$v07 ~ t) )
```
# coefficients of determination =  R square = 0.7103
# At 5%, test for significant of slope: Reject H0: significant

```{r}
# log-linear model
log_linear = lm(log(df$v07) ~ t)
summary( log_linear)
```
```{r}
(exp(log_linear$coefficients[[2]]) - 1 ) *100
```
```{r}
summary(lm( df$v07 ~ t + I(t^2)))
```
đỉnh nằm bên phải
```{r}
length(df$v07)
```
```{r}
d <- c(rep(0,48), rep(1,144))
```
```{r}
summary(lm(df$v07~ t + d))
# bỏ qua tính xu thế, xu hướng giảm
```
```{r}
# Regress on seasonal dummies
seas <- seasdummy(192, 12)
summary(lm(df$v07 ~ seas ))
# coefficient for January = -29.076
```
```{r}
# Regress v07 on time trend and seasonal dummies
# Estimated on difference between Feb and Jan
summary(lm(df$v07 ~ t + seas ))
# bỏ ra tính xu thế, Jan Feb chênh lệch nhau: seas2 - seas1 + t = 30.385
# bỏ ra tính xu thế, Jan March chênh lệch nhau: seas3 - seas1 + 2 * t = 6.65187
```
```{r}
# Regress time trend
regl <- lm(log(df$v07) ~ t + seas)
# First residual
residuals(regl)
```
```{r}
v07 <- ts(df$v07, start = c(2004,1), frequency= 12)
decom.v07.a <- decompose(v07, type = 'additive')
```
```{r}
decom.v07.a$trend
```
```{r}
decom.v07.a$seasonal
```
```{r}
# Random (irregular) component of 2006 January
decom.v07.a$random
```
```{r}
summary(ur.df(v07, type = 'trend', lag= 0))
# coefficient of trend : 0.287
# test for trend should be added: Reject H0, nesscessary
#

# unit root
# H0: unit root
# H1: no
# |tau_stat| < |tau_3_0.05| -> Not Reject H0 -> Unit Root ( + have trend)
# Non-stationary + around trend -> Not trend-stationary
# At 10% reject H0 -> No UR (+ trend) -> stationary around trend -> trend stationary
```
```{r}
summary(ur.df(diff(v07), type = 'trend'))
# At 5%, T-test, trend is nescessary
```
```{r}
summary(ur.df(diff(v07), type = 'drift'))
# At 5%, T-test, const is nescessary
```

```{r}
summary(ur.df(diff(v07), type = 'none'))
```
```{r}
pacf(diff(v07), pl = FALSE)
```
```{r}
acf(diff(v07), plot = FALSE)
```
```{r}
ar1.ls <- arma(diff(v07), order = c(1,0))
summary(ar1.ls)
# long run mean: 3.716 / (1 - (-0.36703))
```
```{r}
ar1m.ls <- arima(diff(v07), order = c(1,0,0))
summary(ar1m.ls)
# Long-run mean intercept: 26889
# variance of shock : -946.5
```
```{r}
summary(arima(v07, order = c(1,1,0)))
```
```{r}
summary(Arima(v07, order = c(1,1,0), include.constant = TRUE))
```
```{r}
summary(Arima(v07, order = c(1,1,0), include.drift = TRUE))
```
```{r}
arima.ls112 <- Arima(v07, order = c(1,1,2), include.constant = TRUE)
summary(arima.ls112)
```
```{r}
checkresiduals(arima.ls112)
# Reject H0 : Residual are Not white noise
```
```{r}
resid <- resid(arima.ls112)
garch11_112 <- garch(resid, order = c(1,1))
```
```{r}
# test for cointegratation (Johansen-procedure)
data_1 = data.frame(df$v05, df$v06, df$v07)
summary(ca.jo(data_1, type = 'trace'))
```
Now, click the **Run** button on the chunk toolbar to [execute](https://www.jetbrains.com/help/pycharm/r-markdown.html#run-r-code) the chunk code. The result should be placed under the chunk.
Click the **Knit and Open Document** to build and preview an output.
